---
title: "isolation_by_distance"
author: "Matthew Bootsma"
date: "April 18, 2019"
output: html_document
---
Running an isolation by distance analysis. Got the first section that uses the ".dist" function from Peter, but the outlier at ~5,0.03 was odd, as was the slope ~=0

I used the matrix of distances I built in GIS using "Point Distance" and the slope got much more biologically relevant. Not sure the unit, the points were in decimal degrees to start so perhaps it's a measure in decimal degrees? At this scale the margin of error due to a 2d vs 3d measure is proportionally small relative to the overall distances so I'm not too worried about making sure it's accounting for curvatrure of the earth.

Scale is correct, again, not sure on units, hopefully I don't have to get this to Miles for the publication but more likely than not I will. Just go ask the GIS department how to measure my points on the map at that point I guess?

#dependencies and genepop
```{r}
library(ggplot2)
library(poppr)
library(adegenet)

total_dat <-read.genepop("../SNPS/v6_SNPs_natural_pops.gen", ncode = 3) 
```
#calculate distances

NOTE: MAKE SURE the lat lon list is ordered the same as your genepop. It looks ok right now but IDK if it's perfect, that outlier on the bottom...
I double checked it and it looks ok... maybe the R calculation is just not as good as the GIS one
The simulation is still usefull to confirm that we have some clear IBD here
```{r}
total_dat_genpop <- genind2genpop(total_dat)

coords <- read.csv("./1710_lat_lon_natural_pops_same_order_as_genpop.csv") # coords = 2 column DF of decimal degrees

geo_dist_total <- dist(coords, method = "euclidean") #the "dist"function then supposedly estimates relative distances (Euclidean).
gen_dist_total <- dist.genpop(total_dat_genpop, method = 2 ) ## method = 2 is Nei's
gen_dist_total
unique(total_dat$pop)
geo_dist_total
```
#mantel test and plot
```{r}
## actual mantel test
ibd_total <- mantel.randtest(gen_dist_total, geo_dist_total)
plot(ibd_total)

plot(geo_dist_total, gen_dist_total, main = "All sites")
abline(lm(gen_dist_total~geo_dist_total), col = "red")
```
#ggplot
```{r}
# #can we color points by pop?
# unique(total_dat$pop)
# color_cat = NULL
# r = 19
# for (i in 1:20) {
#   
#   color_cat = c(color_cat,rep(as.character(unique(total_dat$pop))[i],r))
#   
#   r = r-1
# }
# color_cat 
# #format for ggplot
# v1 = as.numeric(geo_dist_total)
# v1.01 = as.dist(geo_dist_total)
# 
# v2 = as.numeric(gen_dist_total)
# dat_gplot = cbind.data.frame(v1,v2,color_cat)
# names(dat_gplot) = c("geo_dist", "gen_dist","grpz")
# 
# 
# 
# #plot in gg
# ggplot(dat_gplot, aes(geo_dist, gen_dist)) + 
#     geom_point()+geom_smooth(method = "lm", color = "#000099")+ 
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
# 
# #plot in gg
# ggplot(dat_gplot, aes(geo_dist, gen_dist, color = color_cat)) + 
#     geom_point()+ 
#     facet_wrap(color_cat)
#     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
# 
# 
# geo_dist_total[which(gen_dist_total == min(gen_dist_total))]
# names(gen_dist_total)[111]
# geo_dist_total[111]
```
#making key to build my own list of geo_dist calculated in GIS
  I'm not sure the .dist function is working correctly
```{r}
# ID_vec = c(11,17,10,12,1,13,15,2,18,19,3,4,5,6,7,8,9,21,22,23)
# 
# pwise_comparisons = as.data.frame(matrix(nrow = 190, ncol = 2))
# 
# 
# pcomp_current_iter = 2
# for (i in 1:190) {
# 
#   
#   pwise_comparisons[i,1] = ID_vec[1]
#   
#   pwise_comparisons[i,2] = ID_vec[pcomp_current_iter]
#   
#   pcomp_current_iter = pcomp_current_iter+1
#   
#   if (pcomp_current_iter > length(ID_vec)) {
#     ID_vec = ID_vec[-1]
#     pcomp_current_iter = 2
#   }
#   
# }
# 
# write.csv(pwise_comparisons,"./pairwise_comparison_key_for_GIS_geoDist.csv")
```
#plot with my calculated geo_dist
```{r}
my_geo_dist = read.csv("./pairwise_comparison_key_for_GIS_geoDist.csv")

#format for ggplot
v2.2 = as.numeric(gen_dist_total)
dat_gplot2 = cbind.data.frame(my_geo_dist$geo_dist,v2.2)
names(dat_gplot2) = c("geo_dist", "gen_dist")
```


```{r}
pdf("./IBD_v6_natural_pops.pdf")
ggplot(dat_gplot2, aes(geo_dist, gen_dist)) + 
    geom_point()+geom_smooth(method = "lm", color = "#000099")+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  scale_x_continuous(name="Geographic Distance
(Euclidean)", limits=c(0, 10)) +
  scale_y_continuous(name="Genetic Distance
(Edwards)", limits=c(0, 0.2))
dev.off()
```

